---
title: 'Handle your references through `DOI` <br> [<i class="fa fa-github" aria-hidden="true"></i>](https://github.com/SteveViss/Talk_meta_analysis)'
date: "`r paste0(format(Sys.time(), '%d %B %Y'))`"
author: 'Steve Vissault and Kevin Cazelles'
# lang: french
output:
    ioslides_presentation:
        incremental: true
        widescreen: true
        includes:
          in_header: ../assets/kcheader.html  
        #logo: ./img/logo.png
---



```{r dsetup, results = "hide", echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.align='center',
  tidy=FALSE,
  comment="R>> ",
  size = 'small'
)
```

## Briefly, What's a DOI?

- DOI for `Digital object identifier`
- _Unique identifier_ for digital objects (movies to scientific articles)
- At publication stage, each publishers has the responsability to ask for a DOI
- `ISO 26324:2012`

# Retrieve data from DOI

## How can I take advantage of DOI?

[crossref](https://www.crossref.org/) offers web service to retrieve metadata from DOI.

[rcrossref](https://github.com/ropensci/rcrossref) is a R package based on this web service.


## Retrieve data

### Install the package

```r
install.packages('rcrossref')
```

```{r}
library('rcrossref')
vecDOI <- cr_r(3)
vecDOI[1]
```

## Get the bibtex from DOI

### `Bibtex` is a suitable format (supported by Mendeley) which contains all the metadata of an article.

```{r}
bib1 <- cr_cn(vecDOI[1], format='bibentry')
str(bib1)
```

Other format are accessible: TODO

## Get the bibtex from DOI

### Looping of the DOI vector

```{r}
lsf <- lapply(vecDOI, cr_cn, format="bibentry")
nlevels(lsf)
```

## Access the metadata

### Get the author of each articles

```{r}
lsaut <- lapply(lsf, function(x) x$author)
```

### Get the journal of each articles

```{r}
lsjnl <- lapply(lsf, function(x) x$journal)
```

## Perform queries

```{r ex4}
unlist(lsjnl)
```

## Export to csv format

## Export to bibtex format

## Bibliometric studies

<!-- Mettre figure du papier CCvsBL -->
