---
title: 'Importing and Merging spreadsheets <br> [<i class="fa fa-github" aria-hidden="true"></i>](https://github.com/SteveViss/Talk_bib)'
date: "`r paste0(format(Sys.time(), '%d %B %Y'))`"
author: 'Steve Vissault and Kevin Cazelles'
# lang: french
output:
    ioslides_presentation:
        incremental: false
        widescreen: true
        includes:
          in_header: ../assets/kcheader.html  
        #logo: ./img/logo.png
---


```{r dsetup, results = "hide", echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.align='center',
  tidy=FALSE,
  comment="R>> ",
  size = 'small'
)
```

## Two potential work scenarios

**Scenario 1.**

All collaborators work on the _same online collaborative spreadsheet_ (Google sheets).
We want to upload the google spreadsheet in R to perform the analysis.

**Scenario 2.**

Each collaborator works on _different spreadsheet_ with the same format (template) and conventions (units).
We want to merge all csv file into one final csv file to perform the analysis on the entire dataset.

# Scenario 1. <br> All collaborators work on _google spreads_

## Scenario 1. All collaborators work on _google spreads_

### Login and list available worksheets

```{r eval=TRUE}
# load package
library(googlesheets)

# List available google sheets
# Ask you to authenticate in a browser
gs_ls()
```

## Scenario 1. All collaborators work on _google spreads_

### Get your worksheet

```{r}
worksheet <- gs_title("Meta_analyses_google")
```

### List the sheets inside the worksheet

```{r}
gs_ws_ls(worksheet)
```

## Scenario 1. All collaborators work on _google spreads_

### Read the sheet needed to perform the analysis

```{r}
data <- gs_read(ss=worksheet, ws = "Sheet1", verbose=FALSE)
```

## Scenario 1. All collaborators work on _google spreads_

### Transform the `dplyr` object as a conventional dataset object

```{r}
(data <- as.data.frame(data))
```

### And that's it, you're all set!

## Scenario 1. All collaborators work on _google spreads_

### Advantages:

- Secure (dataset stays on the private domain)
- Everyone can work on the same worksheet and avoid conflicts
- Individual sheet are easily accessible

## Scenario 2.

### List the sheets inside the google doc

### Import CSV files

```r
data <- read.csv('file.csv')
```

### Import `.xlsx` files (requires java `sudo R CMD javareconf`):

```r
install.packages('xlsx')
data <- read.xlsx('file.csv')
```

## First scenario: Each collaborator works on different spreadsheet

### Import CSV files

```r
data <- read.csv('file.csv')
```

### Import `.xlsx` files (requires java `sudo R CMD javareconf`):

```r
install.packages('xlsx')
data <- read.xlsx('file.csv')
```



## Multiple spreadsheets

- Use the same headers, put them in a folder then:

```r
fcsv <- list.files('path/to/folder', pattern='^F.*csv$', full.names=TRUE)

lscsv <- list()

for (i in 1:length(fcsv)){
  lscsv <- read.csv(fcsv[[i]])
}
mydat <- do.call('rbind', lscsv)
```

## Multiple spreadsheets


```bash
ssconvert -S fileIn.xlsx foldercsv/fileOut_%s.csv
```

# Google spreadsheet

## Go online

Show how to export them as `.csv` files.



# Alternatives (may be may not be)

## YAML

## JSON

## SQL












## Centered

<div class="centered">
This text is centered.
</div>


## Example with 2 columns

<div class="columns-2">

  - To make it look good
  - I changed the width of the figure  

  ```{r codechunk3, echo=F, fig.width=5.5}
  seqx <- seq(0,10,0.1)
  plot(seqx, cos(seqx), type='h')
  ```

</div>
