---
title: 'Importing and Merging spreadsheets <br> [<i class="fa fa-github" aria-hidden="true"></i>](https://github.com/SteveViss/Talk_meta_analysis)'
date: "`r paste0(format(Sys.time(), '%d %B %Y'))`"
author: 'Steve Vissault and Kevin Cazelles'
# lang: french
output:
    ioslides_presentation:
        incremental: false
        widescreen: true
        includes:
          in_header: ../assets/kcheader.html  
        #logo: ./img/logo.png
---


```{r dsetup, results = "hide", echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  fig.align='center',
  tidy=FALSE,
  comment="R>> ",
  size = 'small'
)
```

## Two potential work scenarios

**Scenario 1.**

All collaborators work on the _same online collaborative spreadsheet_ (Google sheets).
We want to upload the google spreadsheet into R to perform the analysis.

**Scenario 2.**

Each collaborator works on _different spreadsheet_ with the same format (template) and conventions (units).
We want to merge all csv file into one final csv file to perform the analysis on the entire dataset.

# Scenario 1. <br> All collaborators work on _google spreads_

## Scenario 1. All collaborators work on _google spreads_

### Login and list available worksheets

```{r eval=TRUE}
# load package
library(googlesheets)

# List available google sheets
# Ask you to authenticate in a browser
gs_ls()
```

## Scenario 1. All collaborators work on _google spreads_

### Get your worksheet

```{r}
worksheet <- gs_title("Meta_analyses_google")
```

### List the sheets inside the worksheet

```{r}
gs_ws_ls(worksheet)
```

## Scenario 1. All collaborators work on _google spreads_

### Read the sheet needed to perform the analysis

```{r}
data <- gs_read(ss=worksheet, ws = "Sheet1", verbose=FALSE)
```

## Scenario 1. All collaborators work on _google spreads_

### Transform the `dplyr` object as a conventional dataset object

```{r}
(data <- as.data.frame(data))
```

### And that's it, you're all set!

## Scenario 1. All collaborators work on _google spreads_

### Advantages:

- Secure (dataset stays on the private domain)
- Everyone can work on the same worksheet and avoid conflicts
- Individual sheets are easily accessible

# Scenario 2. Each collaborator works on a _different spreadsheet_

## Scenario 2. Each collaborator works on a _different spreadsheet_

### **Note: For this scenario, I assume that headers are consistent among files.**

### Import a flat file (csv format for instance)

```r
data <- read.csv('file.csv')
```

### Import `.xlsx` files (requires java `sudo R CMD javareconf`):

```r
install.packages('xlsx')
data <- read.xlsx('file.csv')
```

## Scenario 2. Each collaborator works on a _different spreadsheet_

### With 3 to 4 collaborators...

```r
data1 <- read.csv('file1.csv')
data2 <- read.csv('file2.csv')
data3 <- read.csv('file3.csv')
data4 <- read.csv('file4.csv')
```

### But... if the team is 15 members (15 files)... you cannot

> ### **Solution**: You have to loop over the files

## Multiple spreadsheets

### Use the same headers, put all of them in a folder then...

```r
# create a vector containing the paths to all files
files <- list.files('path/to/folder', pattern='^F.*csv$', full.names=TRUE)

# Create a empty list
ls_csv <- list()

# The loop populates each level of the list with the content of each file
for (i in 1:length(files)){
  print(files[i])
  ls_csv[[i]] <- read.csv(files[i])
}

# Finally, we apply the function rbind at each level of the list
data <- do.call('rbind', ls_csv)
```

# Exercise

1. Create several fake files (up to 3 CSV files) with common headers
2. Save them in a same folder
3. Open a R script
4. Write down a function merging them (use the previous slide to help)

**Note: Don't hesitate.. if you have any questions, we're here to help you!**
